---
title: "Loops"
output: html_document
date: "2024-10-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# FOR LOOPS

Write a for loop that iterates over the numbers 1 to 5 and prints the cube of each number. Hint to calculate cube: run `5^3` in your Console

```{r}

```


Write a for loop that calculates the MAXIMUM value in each column of diamonds. Preallocate the output as we saw in class. Hint: use `max()`

```{r}
library(tidyverse)
data(diamonds)
diamonds
```

```{r}

```


# MAP()

Write a `map()` function to calculate the MAXIMUM value in each column of diamonds
```{r}

```


Write a `map()` function to calculate the MEAN for every column in diamonds. Remove missing values from your calculation
```{r}

```


# ACROSS()

Use `across()` with `summarize()` to calculate the MAXIMUM value for every column of diamonds
```{r}

```


Use `across()` with `summarize()` and `group_by()` to calculate the MEAN for every NUMERIC column in diamonds for each""cut"
```{r}

```


Calculate the COUNT for *every* factor column in the diamonds dataset
Hint: use `where` https://dplyr.tidyverse.org/dev/articles/colwise.html#multiple-functions
```{r}

```


# REWRITE THE FOLLOWING CODE FROM THE PREVIOUS LECTURE

The code below comes from the in-class exercises on the lecture on conditional statements. Rewrite it by using a for loop instead than `rowwise()`. 

Recall that if / else control flow statements (which are different than `ifelse()` or `if_else()` which are functions) are not vectorized, meaning they evaluate a given condition one element at a time! 
Thus to apply such statements to a dataframe, like in the code below, we need to make sure we force R to process each row individually (e.g., to consider each individual row as its own group). We can achieve this using functions developed for the purpose (like `rowwise()`) or using a for loop. 

Your goal now is to rewrite this code using a for loop instead than `rowwise()`. This is a challanging task: give your best attempt without using AI and we ll then go over the solutions together!

```{r}
# load data first
library(palmerpenguins)
data(penguins)
penguins
```

```{r}
# code to rewrite

penguins %>%
  rowwise() %>%
  mutate(
    mass_category = if (is.na(body_mass_g)) {
      NA_character_
    } else if (body_mass_g > 5000) {
      "Large"
    } else if (body_mass_g >= 4000 & body_mass_g <= 5000) {
      "Medium"
    } else {
      "Small"
    }
  ) %>%
  ungroup() %>%
  group_by(mass_category) %>%
  summarize(
    avg_flipper_length = mean(flipper_length_mm, na.rm = TRUE)
  )
```

```{r}
# code rewritten with a for loop 


```
